Machine Learning and Constraint Programming for Relational-To-Ontology Schema Mapping

### 2018-04-25 stp/paper_aaai/main.tex
gottena  bit of space left

The problem of integrating heterogeneous data sources into an ontology is highly relevant in the database field.
Several techniques exist to approach the problem, but side constraints on the data cannot be easily implemented and thus the results may be inconsistent.
In this paper we improve previous work by \citeasnoun{taheriyan2016learning} using Machine Learning (ML) to take into account inconsistencies in the data (unmatchable attributes) and encode the problem as a Steiner Tree, for which we use work by \citeasnoun{deuna2016steiner} in Constraint Programming (CP).
Combining ML and CP achieves state-of-the-art precision, recall and speed, and provides a more flexible framework for variations of the problem.

@@

Following the traditional data integration framework~\cite{doan2012principles}, we decompose the problem into two parts: schema matching and schema mapping.
Schema matching, which we refer to as \emph{semantic labeling}, finds 
correspondences between attributes from data sources and data nodes of the 
target ontology.
In the schema mapping part we want to generate the semantic models of data sources by identifying the connecting paths for the matched data nodes.

An \emph{attribute mapping} function $\phi : \mathcal{A}_s \mapsto 
\mathcal{D}_m$ is a function which maps the attributes of the source $s$ into the nodes of the semantic model $m$. 
It can be a partial mapping, meaning that only some of the attributes
are connected to the nodes of $m$.
The attribute mapping function addresses the first part of the problem (i.e. the semantic labeling).

@@

We formulate the problem of semantic labeling as a multi-class classification
problem.
The known source descriptions $\Delta_T$ provide us the training sample.
We compute a feature vector for each attribute in a data source and associate 
the known semantic type with the corresponding feature vector.
The feature vector includes, among others, characteristics
such as a number 
of whitespaces and other special characters, statistics of 
values in the column\footnote{Full list of features: https://git.io/vN92f (anonymously hosted)}. 
\forijcai{A full list can be found in Appendix \ref{Ann:SemLab}.}
%\npr{If I remove entropy descr, we are perfectly within page limit}
%\ignore{
One of the important features characterising information content of an attribute is Shannon's entropy.
% of the attribute's concatenated rows
Shannon's entropy 
%(or information entropy~\cite{Manning:Introduction})
 of a string $X$ is defined as
$H(X) = -\sum_{i}{p_i \log_{2}p_i},$ where $p_i$ is the probability of a character, whose index in character vocabulary is $i$, to appear in $X$, and the summation ranges over all
characters in the vocabulary.
%of all possible characters
To evaluate $p_i$ in Shannon's entropy, we evaluate normalized character frequency distribution \emph{chardist} of an attribute, as character counts in concatenated rows of the attribute, normalized by the total length of the concatenated rows.
The alphabet\forijcai{vocabulary of all characters} consists of 100 printable characters\forijcai{ (including $\backslash$n)}.
Finally, we add the 100-dimensional vector of $p_i$ to the attribute feature vector.
%}
We also compute a set of features based on similarity metrics inspired by state-of-the-art works by ~\citeasnoun{Pham:semantic} or~\citeasnoun{Ritze:matching}.
Among others, we compute mean cosine similarity for character distributions of attribute values and string similarity metrics for attribute names.
%To extract features from attribute names, we compute string similarity metrics: minimum edit distance, two WordNet based similarity measures such as JCN~\cite{Jiang:Semantic} and LIN ~\cite{Lin:Information}, and $k$-nearest neighbors using Needle-Wunsch distance~\cite{Needleman:General}.
%The minimum edit distance between two strings $s_1$ and $s_2$ is the minimum number of edit operations, such as insertion, deletion, substitution, which are required to transform one string into another~\cite{Manning:Introduction}.
%We compute the similarity between attribute name and all class instances in the training data.
%The number of thus extracted features depends on the number of semantic labels in the training data.
We train a random forest on the obtained sample.

@@

To evaluate our new system \serene{}, we show that its semantic labeling model 
produces more accurate matches for attributes and that the CP formulation leads 
to better semantic models in terms of precision and recall.

>>>>

The problem of integrating heterogeneous data sources into an ontology is highly relevant in the database field.
Several techniques exist to approach the problem, but side constraints on the data cannot be easily implemented and thus the results may be inconsistent.
In this paper we improve previous work by \citeasnoun{taheriyan2016learning} using Machine Learning (ML) to take into account inconsistencies in the data (unmatchable attributes) and encode the problem as a variation of the Steiner Tree, for which we use work by \citeasnoun{deuna2016steiner} in Constraint Programming (CP).
Combining ML and CP achieves state-of-the-art precision, recall and speed, and provides a more flexible framework for variations of the problem.

@@

Following the traditional data integration framework~\cite{doan2012principles}, we decompose the problem into two parts.
%: schema matching and schema mapping.
First, \emph{semantic labeling} finds 
correspondences between attributes from data sources and data nodes of the 
target ontology.
Second, in the \emph{schema mapping} part we want to generate the semantic models of data sources by identifying the connecting paths for the matched data nodes.

An \emph{attribute mapping} function $\phi : \mathcal{A}_s \mapsto 
\mathcal{D}_m$ maps the attributes of the source $s$ into the nodes of the semantic model $m$. 
It can be a partial mapping, (i.e. only some of the attributes
are connected to the nodes of $m$).
This function addresses the first part of the problem (i.e. the semantic labeling).

@@

We formulate the problem of semantic labeling as a multi-class classification
problem.
The known source descriptions $\Delta_T$ provide us the training sample.
We compute a feature vector for each attribute in a data source and associate 
the known semantic type with the corresponding feature vector.
The feature vector includes, among others, characteristics
such as a number 
of whitespaces and other special characters, statistics of 
values in the column. 
\forijcai{A full list can be found in Appendix \ref{Ann:SemLab}.}
%\npr{If I remove entropy descr, we are perfectly within page limit}
%\ignore{
One of the important features characterising information content of an attribute is Shannon's entropy.
% of the attribute's concatenated rows
Shannon's entropy 
%(or information entropy~\cite{Manning:Introduction})
 of a string $X$ is defined as
$H(X) = -\sum_{i}{p_i \log_{2}p_i},$ where $p_i$ is the probability of a character, whose index in character vocabulary is $i$, to appear in $X$, and the summation ranges over all
characters in the vocabulary.
%of all possible characters
To evaluate $p_i$ in Shannon's entropy, we evaluate normalized character frequency distribution \emph{chardist} of an attribute, as character counts in concatenated rows of the attribute, normalized by the total length of the concatenated rows.
The alphabet\forijcai{vocabulary of all characters} consists of 100 printable characters\forijcai{ (including $\backslash$n)}.
Finally, we add the 100-dimensional vector of $p_i$ to the attribute feature vector.
%}
We also compute a set of features based on similarity metrics inspired by state-of-the-art works by ~\citeasnoun{Pham:semantic} or~\citeasnoun{Ritze:matching}.
Among others, we compute mean cosine similarity for character distributions of attribute values and string similarity metrics for attribute names.
%To extract features from attribute names, we compute string similarity metrics: minimum edit distance, two WordNet based similarity measures such as JCN~\cite{Jiang:Semantic} and LIN ~\cite{Lin:Information}, and $k$-nearest neighbors using Needle-Wunsch distance~\cite{Needleman:General}.
%The minimum edit distance between two strings $s_1$ and $s_2$ is the minimum number of edit operations, such as insertion, deletion, substitution, which are required to transform one string into another~\cite{Manning:Introduction}.
%We compute the similarity between attribute name and all class instances in the training data.
%The number of thus extracted features depends on the number of semantic labels in the training data.
We train a random forest on the obtained sample.

@@

To evaluate our new system \serene{}\footnote{All code and full list of features available at \url{http://github.com/NICTA/serene-benchmark}}, we show that its semantic labeling model 
produces more accurate matches for attributes and that the CP formulation leads 
to better semantic models in terms of precision and recall.
